---
title: "Stat448 Presentation Code"
author: "Lida Tetyusheva"
date: "2023-02-23"
output:
  html_document:
    df_print: paged
  pdf_document: default
subtitle: Urbana Crime Dataset
---

## Project Summary

Using the Urbana Crime Dataset, I am analyzing how crime on Green Street compares to crime in all of Urbana as well as looking for patterns in crime on Green Street.

## Install/Load Packages

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
```

## Import Dataset and Combine

### Urbana Crime Dataset Part 1 and Part 6

```{r}
urbana_part01 <- read_csv("C:/Users/Ivan/Desktop/Stat448Prezi/data_urbana_policing-main/urbana-police-incidents-data-part01.csv", show_col_types = FALSE)

urbana_part06 <- read_csv("C:/Users/Ivan/Desktop/Stat448Prezi/data_urbana_policing-main/urbana-police-incidents-data-part06.csv",show_col_types = FALSE)

urbana = urbana_part01 %>%
  rbind(urbana_part06) %>%
  filter(`YEAR OCCURRED` >= 2013) %>%
  filter_at(vars(`MAPPING ADDRESS`),all_vars(!is.na(.))) %>%
  select(-c(INCIDENT,`DATE REPORTED`,`TIME REPORTED`,`DATE ARRIVED`,`TIME ARRIVED`,STREET,`CROSS STREET`,`CRIME CODE`,`CRIME CATEGORY`,`BIAS CODE`,`PLACE CODE`,`STATUS DESCRIPTION`,`BIAS DESCRIPTION`,`STATUS CODE`, COUNTER,`WEAPONS CODE 2`,`WEAPON 2 DESCRIPTION`,`WEAPONS CODE 3`,`WEAPON 3 DESCRIPTION`,`CSA DESCRIPTION`,`GEO CODE`,`HOUSE NUMBER BLOCK`))%>%
  mutate(`AREA` = case_when(
    `PLACE CODE DESCRIPTION` %in% c("APARTMENT","RESIDENCE-PRIVATE","RESIDENCE-PORCH","DRIVEWAY-RESIDENTIAL","RESIDENCE-YARD") ~ 'Private',
    !(`PLACE CODE DESCRIPTION` %in% c("APARTMENT","RESIDENCE-PRIVATE","RESIDENCE-PORCH","DRIVEWAY-RESIDENTIAL","RESIDENCE-YARD")) ~ 'Public')) 
```

"urbana" Data Description:

| Columns | Rows    | Years     |
|---------|---------|-----------|
| 11      | 113,981 | 2013-2022 |

### Separate between Green Street and Urbana Area

```{r}
green = urbana %>% 
  filter(grepl(' GREEN ',`MAPPING ADDRESS`)) %>%
  filter(!grepl("PRAIRIE",`MAPPING ADDRESS`)) %>%
  mutate(`HUB` = 'Green')

urban = urbana[!(urbana$`MAPPING ADDRESS` %in% green$`MAPPING ADDRESS`),] %>%
  mutate(HUB = 'Urbana')

```

"green" Data Description

| Columns | Rows    | Years     |
|---------|---------|-----------|
| 12      | 1,845   | 2013-2022 |

"urban" Data Description

| Columns | Rows    | Years     |
|---------|---------|-----------|
| 12      | 112,136 | 2013-2022 |

## Top Weapons Used in Crimes Involving Weapons 

#### Green Street

```{r}

count = green %>% count(`WEAPON 1 DESCRIPTION`) %>%
  mutate(percentage = n/sum(n)) %>%
  arrange(desc(percentage)) %>% 
  filter(!row_number() %in% c(1)) %>%
  head(n = 10)

count %>% 
  filter(!row_number() %in% c(10)) %>%
  ggplot(aes(x=`WEAPON 1 DESCRIPTION`, y= percentage)) + 
  geom_bar(stat = "identity", color = 'goldenrod2', fill = 'gold')+
  ggtitle("Green Street") +
  coord_flip()

```

#### Urbana 

```{r}
count2 = urban %>% count(`WEAPON 1 DESCRIPTION`) %>%
  mutate(percentage = n/sum(n)) %>% 
  arrange(desc(percentage)) %>% 
  filter(!row_number() %in% c(1)) %>%
  head(n = 10)

count2 %>% 
  filter(!row_number() %in% c(25)) %>%
  ggplot(aes(x=`WEAPON 1 DESCRIPTION`, y= percentage)) + 
  geom_bar(stat = "identity", color = 'goldenrod2', fill = 'gold')+
  ggtitle("Urbana")+
  coord_flip()

```

Observations:
- NONE, PERSONAL WEAPONS, and HANDGUN seem to be the most common weapons used when 
a crime involves a weapon
- unknown what NONE description is referring to, but for the sake of this dataset, 
we will keep it in the analysis.
- percentage of weapons used in crime seems higher in Green Street

## Testing Significance for Difference in Percentage of Weapon Use
#### Using Two Proportions Z-test at 95% confidence

```{r}

prop.test(x = c(sum(count$n),sum(count2$n)), n = c(nrow(green),nrow(urban)), alternative = 'two.sided',correct = TRUE)$p.value

prop.test(x = c(sum(count$n),sum(count2$n)), n = c(nrow(green), nrow(urban)), alternative = 'greater')$p.value

```

Conclusion: 
At $\alpha$ = .05, the percentage of crimes involving weapons on Green Street is **significantly different** and **significantly greater** than that of Urbana.

## Time of Crime vs. Number of Crimes
### Green Street
```{r}
green %>%
  mutate(`TIME OCCURRED` = hour(hms(green$`TIME OCCURRED`))) %>%
  group_by(`HUB`)%>%
  count(`TIME OCCURRED`)%>%
  ggplot(aes(x= `TIME OCCURRED`, y= n)) +
  ggtitle("Hour vs Crime Frequency on Green Street") +
  geom_line(color = 'blue', size = 1) 
```

#### Urbana
```{r}
urban %>%
  mutate(`TIME OCCURRED` = hour(hms(urban$`TIME OCCURRED`))) %>%
  filter(`TIME OCCURRED` <= 23) %>%
  group_by(`HUB`)%>%
  count(`TIME OCCURRED`)%>%
  ggplot(aes(x= `TIME OCCURRED`, y= n)) +
  ggtitle("Hour vs Crime Frequency in Urbana") +
  geom_line(color = 'orange', size = 1) 
```
Observations:
The time a crime occurs seems to peak around 4 o'clock on Green Street and around 12 o'clock in Urbana. If most crimes on Green Street are being committed by campus residents and students, I would guess the peak is at 4 because that's when students tend to end their school day. 

## Number of Crimes vs. the Month
#### Green Street
```{r}
green %>%
  count(`MONTH OCCURRED`)%>%
  ggplot(aes(x = `MONTH OCCURRED`, y = n, fill = `MONTH OCCURRED`)) +
  ggtitle("Green Street Crime Per Month") +
  geom_bar(stat = 'identity' )+
  theme(legend.position="none")
```
#### Urbana
```{r}
urban%>%
  count(`MONTH OCCURRED`)%>%
  ggplot(aes(x = `MONTH OCCURRED`, y = n, fill = `MONTH OCCURRED`)) +
  ggtitle("Urbana Crime Per Month") +
  geom_bar(stat = 'identity' )+
  theme(legend.position="none")
```

Observations:
Urbana appears to have a smooth slope that peaks in August, whereas Green Street seems to have two peaks in April and October. I wonder if in the Green Street data there is a significant relationship between time periods in the year and mean number of crime that relates to the school year. 


